---
title: "r_class_1"
author: "Albert Camps"
date: "27 de marzo de 2015"
output: html_document
runtime: shiny
---

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

## Inputs and Outputs

You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot.

```{r, echo=FALSE}
inputPanel(
  selectInput("n_breaks", label = "Number of bins:",
              choices = c(10, 20, 35, 50), selected = 20),
  
  sliderInput("bw_adjust", label = "Bandwidth adjustment:",
              min = 0.2, max = 2, value = 1, step = 0.2)
)

renderPlot({
  hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks),
       xlab = "Duration (minutes)", main = "Geyser eruption duration")
  
  dens <- density(faithful$eruptions, adjust = input$bw_adjust)
  lines(dens, col = "blue")
})
```

## Embedded Application

It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory:

```{r, echo=FALSE}
shinyAppDir(
  system.file("examples/06_tabsets", package="shiny"),
  options=list(
    width="100%", height=550
  )
)
```

Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy.

You can also use the `shinyApp` function to define an application inline rather then in an external directory.

In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components.

----
#Session 1 of R training
Preferences -> panes

Assignations
``` {r}
a <- 5
```
echo=FALSE means that we don't want to see the code, only the output.

There are default packages with data to _play_ around. Like `cars`.

[R markdown](http://markdown.rstudio.com)

## Working with data.

### Types

- Numbers: logicals(0/1), integer, numeric or double.
- Strings (type character) -- single and double quotations have the same meaning
- Dates

## Homogeneous structures

All elements of structure have the same type.

- Vector

```{r}
a <- c(3,4,5)
a[2]
class(a)

s <- c("a", "b", "c")
class(s)

# Convenient ways to create vectors
1:5
seq(0,100,10)
rep("char", 3)

# Vector operations
3*a

# If you have different types, it will try to mix them into the same type
c(1,2, c(1,2))
c(1,2,"b")
```
You can run commands with ctrl+<Intro>

```{r}
mat <- matrix(c(a,a,a), ncol=3)
t(mat) # transpose
mat[1,2]
mat[1,]
mat[,3]
```
alt + cmd + i creates a chunk for markdown

- arrays (n-dimensional)

### Heterogeneous structures

Can contain elements of different types

- lists:

```{r}
myList <- list(
  country = c("Portugal", "Italy", "Spain", "Greece"),
  GDPgrowth = c(0.1,0.5,2,-1.5)
)
```

Use dollar sign (%) to call elements in lists:

```{r}
myList$country
# to know the names of the elements of a list:
names(myList)
summary(myList)
summary(myList$country)
```

- Data frames
```{r}
myDF <- data.frame(
  country = c("Portugal", "Italy", "Spain", "Greece"),
  GDPgrowth = c(0.1,0.5,2,-1.5),
  k = 3
)

#selecting elements of a dataframe
myDF[1,]
myDF[,3]
myDF[,"GDPgrowth"]
myDF[,c("GDPgrowth","k")]
# We can change the names of the columns
names(myDF)
names(myDF) <- c("coun", "gpd", "g")

# If you want just a column you can use the name notation
myDF$gpd

#subsetting
positive_g <- myDF$gpd>0
myDF[positive_g,]
myDF[myDF$coun == "Spain",]
myDF

# Recycle condition for selecting rows when it does not have the ssame length as data.frame

alternate <- c(TRUE, FALSE)
myDF[alternate,]

# Useful functions for dfs:
nrow(myDF)
ncol(myDF)
dim(myDF)

# To look at the first or last rows of my data frame
head(myDF,2)
tail(myDF,2)

View(myDF)
summary(myDF)
```

The `with` function can be quite convenient:

```{r}
# Instead of
sum(iris$Sepal.Length*iris$Sepal.Width/sum(iris$Petal.Length*iris$Petal.Width))
# we can write
with(iris, sum(Sepal.Length*Sepal.Width/sum(Peta.Length*Petal.Width)))
```

### Factor variables
Categorical variables, displayed a name but internally it is saved as a integer.

```{r}
# Create new variable in df
mtcars2 <- mtcars
mtcars2$hp_cat <- "High"
mtcars2[mtcars2$hp < mean(mtcars2$hp),]$hp_cat <- "Low"
mtcars2$hp_factor <- factor(mtcars2$hp_cat)

levels(mtcars2$hp_factor)
mtcars2["Datsun 710", ]$hp_factor <- "Mid"
mtcars2["Datsun 710",]$hp_cat <- "Mid"
```

Beware of factors, since some times you won't want your imported strings to become factors [Default behaviour in R]

### Converting between types

```{r}
mtcars2$cyl_factor <- as.factor(mtcars2$cyl)
mtcars2$cyl_factor
levels(mtcars2$cyl_factor)
mtcars2$mpg_str <- as.character(mtcars2$mpg)
#View(mtcars2)
is.factor(mtcars2$cyl)
is.factor(mtcars2$cyl_factor)
```

### Manipulating strings

concatenating strings with `paste`:

```{r}
paste("Social", "Point")
paste("Social", "Point", sep="")
sp <- paste("Social", "Point", sep=".")
paste(month.abb, collapse="/")
```
In your chunk you can have error=TRUE to display the error
```{r}
nchar(sp)
nchar(month.abb)
toupper(month.abb)
tolower(month.abb)
substr("SocialPoint", 1,6)
# Regular expressions
grepl("Poi", )
gsub("Poi", "XXX", sp)
```

## Dates and times

```{r}
Sys.Date()
as.numeric(Sys.Date())
as.Date("20150323", format="%Y%m%d")
as.Date("12-04-26 16:27:09", format="%y-%m-%d %H:%M:%S")
as.POSIXct("12-04-26 16:27:09", format="%y-%m-%d %H:%M:%S")
# There is also POSIXlt when working with dates
weekdays(Sys.Date())
# With google search for dateformats
```


