---
title: "R-training-session-2"
author: "Albert Camps"
date: "1 de abril de 2015"
output: html_document
runtime: shiny
---

To connect to our database:

```{r}
library(RPostgreSQL)
drv <- dbDriver("PostgreSQL")
spdb <- dbConnect(
   drv,
   host = "socialpoint.com",
   port = "1111",
   dbname = "datawarehouse",
   user = "user",
   password = "*****"
)
dbListTables(spdb)
```

```{r}
jh_user <-dbGetQuery(spdb, "select * from jurassichunter.t_user limit 1000")
jh_session <- dbGetQuery(spdb, "select * from jurassichunter.t_session ")
```

Inspect the data:

```{r}
str(jh_user)
str(jh_session)
```

## Data wrangling

Take a sample of users:

```{r}
#To take a sample, we could either pass a range, or we could pass a logical vector with the same lenght, and true for the ones selected.
# Or we could have a vector indicating which ids we would be using
sample1000 <- sample(1:nrow(jh_user), 1000)
users1000 <- jh_user[sample1000,]
jh_user[,]

# Fraction. We indicate how many we want based on a percentage.
sample2pct <- runif(nrow(jh_user)) <= 0.02
users2pct <- jh_user[sample2pct,]

#Finally we could do the same subsetting.
users2pctSubset <- subset(
  jh_user,
  subset = sample2pct,
  select = c(user_id, register_ip_country, is_tester)
)

sum(sample2pct)
#it is approximatedly the 2%, and usually this will be good enough.
sum(sample2pct/nrow(jh_user))

# let's put them together

#but this will not work, because they have not the same columns
users0 <- rbind(users1000, users2pct)

users0 <- rbind(users1000[, names(users2pct)], users2pct)

```

- Remove duplicates: 

```{r}
dups <- duplicated(users0$user_id)
users1 <- users0[!dups,]
nrow(users0)
nrow(users1)
```

Keep Australians and Spaniards and remove testers

```{r}
str(users0)
spaniards <- users0$register_ip_country == "ES"
australians <- users0$register_ip_country == "AU"
not_testers <- !users0$is_tester
# YOu have it here. NOTE: Logical selectors are single here, not double.
users0[(spaniards | australians) & not_testers,]

# a bit more verbose
sp_users <-users0[spaniards,]
au_users <- users0[australians,]
sp_not_testers <- !sp_users$is_tester
sp_selected <- sp_users[sp_not_testers,]
au_not_testers <- !au_users$is_tester
au_selected <- au_users[au_not_testers,]
total <- c(au_selected, sp_selected)

# or doing a condition straight
es_au_nontest <- (users1$register_ip_country == "ES" | users1$register_ip_country == "AU") & users1$is_tester==FALSE

# yet another solution
es_au_nontest <- which((users1$register_ip_country == "ES" | users1$register_ip_country == "AU") & users1$is_tester==FALSE)
users2 <- subset(subset=es_au_nontest, x=users1)
```

Sessions: keep user_id, start and end date/times, end level. Remove missing values.

```{r}
missingValues <- is.na(jh_session$user_id) |
  is.na(jh_session$date_session_start) | 
  is.na(jh_session$date_session_end) |
  is.na(jh_session$game_end_level)

session0 <- subset(
  jh_session,
  !missingValues,
  select = c(user_id, date_session_start, date_session_end, game_end_level)
  )
```

How to get only the first session of a user.

```{r}
ordered_sessions <- session0[order("user_id","date_session_start"),]
order <- order(session0$date_session_start)
ordered_sessions <- session0[order,]
dups <- duplicated(ordered_sessions$user_id)
cleaned_sessions <- ordered_sessions[!dups,]
sort_vector <- order(session0$user_id, session0$date_session_start)
session1 <- session0[sort_vector,]
session1 <- session1[!duplicated(session1$user_id),]
str(ordered_sessions)
str(cleaned_sessions)
#head(jh_session)
```

Convert to POSIX:

```{r}
library(lubridate)
session1$date_session_start <- ymd_hms(session1$date_session_start)
session1$date_session_end <- ymd_hms(session1$date_session_end)
```

Merge users and sessions

```{r}
#by default it performs an inner join
user_session <- merge(users2, session1, by = "user_id")
```

Session length in hours:

```{r}
user_session$session_length <- difftime(
  user_session$date_session_end,
  user_session$date_session_start,
  units="hours"
  )
# the result is a difftime class, so we would need to convert it to operate
user_session$session_hours <- as.numeric(user_session$session_length)
```

Measure skill:

```{r}
user_session$skill <- user_session$game_end_level / user_session$session_hours

#let's check the average
summary(user_session$skill)
```

User segmentation by skill:
```{r}

```


